---
title: "Key Takeaways from game 3: Celtics v Warriors"
description: |
  Celtics pick and roll defence improves, Klay finds his range
author:
  - name: Mike Toolan
    url: {}
date: 2022-06-09
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<iframe width="560" height="315" src="https://www.youtube.com/embed/lFovcuvoGCg" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

### Stick to the right thing, even at a cost

One of the 'almost' turning points in game 3 was the moment Steph Curry was fouled by Al Horford while making a three.


<iframe width="560" height="315" src="https://www.youtube.com/embed/n0KPBcrD6Rs" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


In an attempt to get contest the shot, Horford rushed towards Curry after his defender had set a screen. Unfortunately for Horford, Curry drilled the three while also getting caught by the unrushing Horford as he made his way back to Earth. After a review, the officials determined it was a flagrant 1 on Horford and that the Warriors would get a free throw and the ball back. This single play was a 7 point swing in favour of the Warriors and nearly resulted in an unlikely comeback, cutting the Celtics lead from 9 to 2.

Here's the thing, _Horford was right_. Curry has shown time and again in this series how prolific he is off ball screens and game 3 was no exception. Any hesitation, gap, or space he can find after receiving a screen is equivalent to a wide open three for most other players. Horford, recognizing this, tried in desperation to shut down the space, giving up the foul. While this was not the desired result, _on average_ the Celtics were very successful in shutting down Curry's attempts from 3 compared to game 1. 

```{r Loading libraries and getting data}
library(nbastatR)
library(tidyverse)
library(BasketballAnalyzeR)
library(jsonlite)
library(janitor)
library(extrafont)
library(ggrepel)
library(scales)
library(teamcolors)
library(zoo)
library(future)
library(lubridate)
library(ggplot2)
library(gridExtra)
library(grid)
library(cowplot)

Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)

warriors_shots = teams_shots(teams = "Golden State Warriors", seasons = 2022, season_types = "Playoffs")


game_1_date = 20220602
game_2_date = 20220605
game_3_date = 20220608

curry_v_boston = warriors_shots %>% filter(., namePlayer == "Stephen Curry",
                                  dateGame == c(game_1_date,
                                               game_2_date,
                                               game_3_date))

curry_season = bref_players_stats(seasons = 2022) %>%
  filter(namePlayer == "Stephen Curry")

reg_season_fga_per_game = curry_season$fgaTotals / curry_season$countGames %>%
  set_names(., "FGAPG")

boston_fga_per_game = round(nrow(curry_v_boston) / 3, 2) %>%
  set_names(., "FGAPG")

ratio = reg_season_fga_per_game / boston_fga_per_game %>%
  set_names("ratio")

curry_shooting_stats = bind_rows(reg_season_fga_per_game, 
                                 boston_fga_per_game) %>%
  mutate(season_stage = c("Regular Season", "NBA Final"))

curry_field_goal_attempts = curry_shooting_stats %>%
  ggplot(., aes(x = season_stage, 
                                   y = FGAPG)) + 
    geom_bar(stat = "identity",  fill="#520f0f") + 
  geom_text(aes(label = FGAPG), vjust=1.6, color="white", size=3.5) +
  theme_minimal() + 
  labs(title="Steph Curry Shots Per Game", 
         x="Stage of Season", y = "Field Goals Attempted") +   
  scale_x_discrete(limits=c("Regular Season", "NBA Final"))

curry_field_goal_attempts

```

```{r Create Court}
circle_points = function(center = c(0, 0), radius = 1, npoints = 360) {
  angles = seq(0, 2 * pi, length.out = npoints)
  return(data_frame(x = center[1] + radius * cos(angles),
                    y = center[2] + radius * sin(angles)))
}
# Court Dimenons & lines
width = 50
height = 94 / 2
key_height = 19
inner_key_width = 12
outer_key_width = 16
backboard_width = 6
backboard_offset = 4
neck_length = 0.5
hoop_radius = 0.75
hoop_center_y = backboard_offset + neck_length + hoop_radius
three_point_radius = 23.75
three_point_side_radius = 22
three_point_side_height = 14
# Court themes
court_themes = list(
  light = list(
    court = 'floralwhite',
    lines = '#999999',
    text = '#222222',
    made = '#00bfc4',
    missed = '#f8766d',
    hex_border_size = 1,
    hex_border_color = "#000000"
  ),
  dark = list(
    court = '#000004',
    lines = '#999999',
    text = '#f0f0f0',
    made = '#00bfc4',
    missed = '#f8766d',
    hex_border_size = 0,
    hex_border_color = "#000000"
  ),
  ppt = list(
    court = 'gray20',
    lines = 'white',
    text = '#f0f0f0',
    made = '#00bfc4',
    missed = '#f8766d',
    hex_border_size = 0,
    hex_border_color = "gray20"
)
)
# Function to create court based on given dimensions
plot_court = function(court_theme = court_themes$light, use_short_three = FALSE) {
  if (use_short_three) {
    three_point_radius = 22
    three_point_side_height = 0
  }
  
  court_points = data_frame(
    x = c(width / 2, width / 2, -width / 2, -width / 2, width / 2),
    y = c(height, 0, 0, height, height),
    desc = "perimeter"
  )
  
  court_points = bind_rows(court_points , data_frame(
    x = c(outer_key_width / 2, outer_key_width / 2, -outer_key_width / 2, -outer_key_width / 2),
    y = c(0, key_height, key_height, 0),
    desc = "outer_key"
  ))
  
  court_points = bind_rows(court_points , data_frame(
    x = c(-backboard_width / 2, backboard_width / 2),
    y = c(backboard_offset, backboard_offset),
    desc = "backboard"
  ))
  
  court_points = bind_rows(court_points , data_frame(
    x = c(0, 0), y = c(backboard_offset, backboard_offset + neck_length), desc = "neck"
  ))
  
  foul_circle = circle_points(center = c(0, key_height), radius = inner_key_width / 2)
  
  foul_circle_top = filter(foul_circle, y > key_height) %>%
    mutate(desc = "foul_circle_top")
  
  foul_circle_bottom = filter(foul_circle, y < key_height) %>%
    mutate(
      angle = atan((y - key_height) / x) * 180 / pi,
      angle_group = floor((angle - 5.625) / 11.25),
      desc = paste0("foul_circle_bottom_", angle_group)
    ) %>%
    filter(angle_group %% 2 == 0) %>%
    select(x, y, desc)
  
  hoop = circle_points(center = c(0, hoop_center_y), radius = hoop_radius) %>%
    mutate(desc = "hoop")
  
  restricted = circle_points(center = c(0, hoop_center_y), radius = 4) %>%
    filter(y >= hoop_center_y) %>%
    mutate(desc = "restricted")
  
  three_point_circle = circle_points(center = c(0, hoop_center_y), radius = three_point_radius) %>%
    filter(y >= three_point_side_height, y >= hoop_center_y)
  
  three_point_line = data_frame(
    x = c(three_point_side_radius, three_point_side_radius, three_point_circle$x, -three_point_side_radius, -three_point_side_radius),
    y = c(0, three_point_side_height, three_point_circle$y, three_point_side_height, 0),
    desc = "three_point_line"
  )
  
  court_points = bind_rows(
    court_points,
    foul_circle_top,
    foul_circle_bottom,
    hoop,
    restricted,
    three_point_line
  )
  
  
  court_points <- court_points
  
  # Final plot creation
  ggplot() +
    geom_path(
      data = court_points,
      aes(x = x, y = y, group = desc),
      color = court_theme$lines
    ) +
    coord_fixed(ylim = c(0, 45), xlim = c(-25, 25)) +
    theme_minimal(base_size = 22) +
    theme(
      text = element_text(color = court_theme$text),
      plot.background = element_rect(fill = 'gray20', color = 'gray20'),
      panel.background = element_rect(fill = court_theme$court, color = court_theme$court),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      axis.text = element_blank(),
      axis.title = element_blank(),
      axis.ticks = element_blank(),
      legend.background = element_rect(fill = court_theme$court, color = court_theme$court),
      legend.margin = margin(-1, 0, 0, 0, unit = "lines"),
      legend.position = "bottom",
      legend.key = element_blank(),
      legend.text = element_text(size = rel(1.0))
    )
}
```

```{r Steph Shot Chart}

curry_chart = curry_v_boston %>%
  mutate(x = as.numeric(as.character(locationX)) / 10, y = as.numeric(as.character(locationY)) / 10 + hoop_center_y)


p1 <- plot_court(court_themes$dark, use_short_three = F) +
  geom_point(data = curry_chart, aes(x = x, y = y, color = curry_v_boston$isShotMade, fill = curry_v_boston$isShotMade), 
             size =3, shape = 21, stroke = .5) +  
  scale_color_manual(values = c("green4","red3"), aesthetics = "color", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_fill_manual(values = c("green2","red3"), aesthetics = "fill", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_x_continuous(limits = c(-27.5, 27.5)) +
  scale_y_continuous(limits = c(0, 45)) +
  theme(plot.title = element_text(hjust = .5, size = 22, family = "Comic Sans MS", face = "bold", vjust = -4),
        plot.subtitle = element_text(hjust = .5, size = 10, family = "Comic Sans MS", face = "bold", vjust = -8),
        legend.position = c(.5, .85),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.text = element_text(hjust = .5, size = 10, family = "Comic Sans MS", face = "bold", colour = "white"),
        plot.caption = element_text(hjust = .5, size = 6, family = "Comic Sans MS", face = "bold", colour = "lightgrey", vjust = 8)) +
  ggtitle(label = "Stephen Curry vs. Boston",
          subtitle = "Games 1 - 3 NBA Finals 2022") 
ggdraw(p1) + theme(plot.background = element_rect(fill="gray20", color = NA)) 


```

```{r Comparison Shot Chart}
#curry_chart_2 = as.data.frame(curry_v_boston)

#PbP <- PbPmanipulation(PbP.BDB)
#subdata <- subset(PbP, player=="Stephen Curry")

#shotchart(data = curry_chart_2, 
 #          x = curry_chart_2$locationX, 
  #         y = curry_chart_2$locationY,
   #        scatter = TRUE)
```